---
title: "CO2 vs Happiness"
author: "Katherien Pryal"
date: "11/29/2020"
output: html_document
---

```{r echo=FALSE, message=FALSE}
library(knitr)
library(dplyr)
```

# CO2 per Capita vs Happiness Table

```{r echo=FALSE, message=FALSE}
source("scripts/summaryTable.R")
table <- make_summary_table(data)
kable(table, col.names = c("Rounded Happiness Score", "Mean CO2 per Capita", "Minimum CO2 per Capita", "Maximum CO2 per Capita"))
```
This table narrows down the comparison between a country's happiness score and 
their CO2 emissions. I have combined countries into groups, bases of of their
happiness score rounded to the closest whole number. In the table above you will 
see that mean CO2 emissions per capita increases as the happiness score 
increases, but this excludes the highest (whole number) happiness score of 8. 
The average CO2 emissions per capita for countries in the group with a score of
8 actually decreases. This could suggest that CO2 emissions increase happiness 
only to a certain point. In the table we also see that the country with the minimum CO2 emissions per capita ranked fairly high in the happiness score with a high score of 5. The country with the max CO2 emissions per capita was not far off with a score of 4. This suggests that there are other factors to consider in comparing CO2 emmistions to happiness. 