---
title: "CO2 vs Happiness"
author: "Katherine Pryal"
date: "11/29/2020"
output: html_document
---

```{r echo=FALSE, message=FALSE}
library(knitr)
library(dplyr)
library(ggplot2)
df <- read.csv("data/final_data.csv", stringsAsFactors = FALSE)
```
# Domain of Interest

Carbon dioxide (CO2) generates the largest share of greenhouse emissions in the 
world. CO2 is released during industrial processes and fossil fuel combustions 
to retrieve energy, or from direct human-induced impacts on forestry and other 
land use. Emissions have been increasing in the atmosphere at an alarming rate, 
impacting our environments and overall quality of life. And recent [studies](https://www.gfmag.com/global-data/non-economic-data/happiest-countries) 
have revealed that the happiest countries are those prioritizing well-being and 
environmental sustainability. To explore this finding, our group is interested 
in understanding the association between CO2 emissions per capita and happiness 
scores by country. We are using a data sets from [The World Happiness Report](https://www.kaggle.com/unsdsn/world-happiness) and [CO2 and Greenhouse 
Gas Emissions by Our World Data](https://github.com/owid/co2-data#%EF%B8%8F-download-our-complete-co2-and-greenhouse-gas-emissions-dataset--csv--xlsx--json).


# Summary Information
```{r, echo=FALSE, message=FALSE, results='hide'}
source("scripts/summary_info.R")
summary <- get_summary_info(df)
summary
```


The joined data set that we are using for this project consists of 2018 
happiness values on a scale of 1-10 from the World Happiness Report and CO2
emissions from 2018 measured in millions of metric tons from 
`r summary$number_of_countries` countries worldwide. In the study, happiness 
scores range from Burundi's `r summary$lowest_happiness` to Finland's 
`r summary$highest_happiness`. On the other hand, there is an even larger range
of CO2 emissions as the lowest country, the Central African Republic, only
produces `r summary$lowest_co2` million tonnes while China produces the most 
with `r summary$highest_co2` million tonnes, meaning China produces 
`r summary$co2_comparison` times more. However, this comparison does not 
factor in how many more people China has than low emission countries like the 
Central African Republic which is why our group's focus is on the CO2 emissions 
per capita data. The range for this is much smaller, from Somalia's
`r summary$lowest_co2_per_capita` million tonnes per capita to Qatar's
`r summary$highest_co2_per_capita` million tonnes per capita, for a difference 
of only `r summary$per_capita_comparison` times more.


# CO2 per Capita vs Happiness Table

```{r, echo=FALSE, message=FALSE}
source("scripts/summary_table.R")
table <- make_summary_table(df)
kable(table, col.names = c("Rounded Happiness Score", "Mean CO2 per Capita", "Minimum CO2 per Capita", "Maximum CO2 per Capita"))
```

### What the chart attempts to seeks to express:
This table narrows down the comparison between a country’s happiness score and 
their CO2 emissions. I have combined countries into groups, bases on their 
happiness score rounded to the closest whole number. I organized this chart in 
this way to show the correlation between happiness and CO2 emissions.


### What information this table reveals: 
In the table above you will see that mean CO2 emissions per capita increases as 
the happiness score 
increases, but this excludes the highest (whole number) happiness score of 8. 
The average CO2 emissions per capita for countries in the group with a score of 
8 actually decreases. This could suggest that CO2 emissions increase happiness 
only to a certain point. In the table we also see that the country with the 
lowest CO2 emissions per capita ranked fairly high in the happiness score with 
a high score of 5. The country with the highest CO2 emissions per capita was not
far off with a score of 4. This suggests that there are other factors to 
consider in comparing CO2 emissions to happiness then what is shown in this 
table.

# Life Expectancy Among with the Countries with the Lowest CO2 Emissions
```{r echo=FALSE, message=FALSE, warning=FALSE}
source("scripts/pi_chart.R")
chart_three <- make_chart_three(df)
chart_three
```


### What the chart attempts to express:
Among the 6 countries with the lowest CO2 emissions, how are the life 
expectancies proportioned to each other? 
Among these countries, are there vast differences in life expectancy or 
are the life expectancies similar? 

### Why we selected chart type and graphical encoding:
We used a pie chart to effectively display healthy life 
expectancy proportionally among the select small group of countries 
with lowest CO2 emissions. The pie chart allows the viewer to see how these 
country's life expectancies are proportional to each other. 

### What information the chart reveals:
The pie chart suggests that there are significant differences in life
expectancy among the the top 6 countries with the lowest CO2 emissions per
capita. For example, Kosovo's life expectancy alone is only slighly less than
the life expetancies of the 5 other countries combined. Thus, life expectancy
varies greatly among the countries with the lowest CO2 per capita emissions and
can not be considered similar. 


# Happiness Scores for the Lowest CO2 Per Capita Countries Bar Chart

```{r echo=FALSE, message=FALSE}
source("scripts/bar_chart.R")
chart_co2_happiness_bar <- co2_happiness_bar(df)
chart_co2_happiness_bar
```

This chart displays the happiness score for each of the five countries with
the lowest CO2 per capita.

### What the chart attempts to seeks to express:
  What are the happiness scores of the 5 countries with 
  the lowest CO2 per capita in 2018? 
  
### Why we selected chart type and graphical encoding:
  We chose a bar chart to answer this question because bar charts easily 
  allow viewers to see the happiness score of each of the 5 countries and how
  the happiness score of these countries compare to the others selected. The
  chart is interactive so the score can be viewed when a bar is selected.

### What information this chart reveals: 
  The happiness scores for the 5 countries with the lowest CO2 per Capita 
  are 2.9 (Burundi), 3 (Central African Republic), 4.3 (Chad), 3.6 (Malawi), 
  and 5 (Somalia). This is interesting because the happiness scores of these 
  countries are in the bottom and middle of the range of happiness scores, thus
  the top five lowest CO2 per Capita countries are not countries with the 
  highest happiness scores.
  
# CO2 Per Capita and Happiness Score Scatter Plot

```{r echo=FALSE, message=FALSE, warning = FALSE}
source("scripts/scatter_plot_chart.R")
chart_co2_happiness_scatter <- co2_happiness_scatter(df)
chart_co2_happiness_scatter
```
This scatter plot displays the relationship between CO2 per capita and
happiness scores for each country. Each dot on the plot represents the CO2
per capita and happiness score for a specific country. The yellow line is a 
smoothing line based on the plot data. 


### What the chart attempts to seeks to express: 
What relationship does the happiness score and CO2 per capita 
by country in 2018 have? Are countries with higher CO2 per capita happier or 
less happy?

### Why we selected chart type and graphical encoding:
We chose a scatter plot because scatter plots display the relation between 
2 variables. The individual points on the scatter plot allow the viewer to
see the happiness score and CO2 per capita for each country. The smoothing line
is also important for answering our question because it depicts the overall
scatter plot trend based on the individual points so that the relation between
these variables is clearly depicted. The scatter plot is interactive so the 
happiness score and CO2 per capita can be viewed when a point is selected.
  
### What information the chart reveals: 
The scatter plot suggests a positive relationship between CO2 per
capita and happiness score by country. Thus, in general, countries with higher
happiness scores also have higher co2 per capita emissions.
