---
title: "CO2 vs Happiness"
author: "Katherine Pryal"
date: "11/29/2020"
output: html_document
---

```{r echo=FALSE, message=FALSE}
library(knitr)
library(dplyr)
library(ggplot2)
df <- read.csv("data/final_data.csv", stringsAsFactors = FALSE)
```
# Domain of Interest

Carbon dioxide (CO2) generates the largest share of greenhouse emissions in the world. CO2 is released during industrial processes and fossil fuel combustions to retrieve energy, or from direct human-induced impacts on forestry and other land use. Emissions have been increasing in the atmosphere at an alarming rate, impacting our environments and overall quality of life. And recent [studies](https://www.gfmag.com/global-data/non-economic-data/happiest-countries) have revealed that the happiest countries are those prioritizing well-being and environmental sustainability. To explore this finding, our group is interested in understanding the association between CO2 emissions per capita and happiness scores by country.

# CO2 per Capita vs Happiness Table

```{r, echo=FALSE, message=FALSE}
<<<<<<< HEAD
source("scripts/summaryTable.R")
table <- make_summary_table(df)
=======
source("scripts/summary_table.R")
table <- make_summary_table(data)
>>>>>>> 719577fee8e2cfe3a6fb4c14fc0800dca3679521
kable(table, col.names = c("Rounded Happiness Score", "Mean CO2 per Capita", "Minimum CO2 per Capita", "Maximum CO2 per Capita"))
```
### What the chart attempts to seeks to express:
This table narrows down the comparison between a country’s happiness score and 
their CO2 emissions. I have combined countries into groups, bases on their 
happiness score rounded to the closest whole number. I organized this chart in 
this way to show the correlation between happiness and CO2 emissions.


### What information this table reveals: 

In the table above you will see that mean CO2 emissions per capita increases as 
the happiness score 
increases, but this excludes the highest (whole number) happiness score of 8. 
The average CO2 emissions per capita for countries in the group with a score of 
8 actually decreases. This could suggest that CO2 emissions increase happiness 
only to a certain point. In the table we also see that the country with the 
lowest CO2 emissions per capita ranked fairly high in the happiness score with 
a high score of 5. The country with the highest CO2 emissions per capita was not
far off with a score of 4. This suggests that there are other factors to 
consider in comparing CO2 emissions to happiness then what is shown in this 
table.

# Comparing Life Expectancy with CO2 Emissions
```{r echo=FALSE, message=FALSE, warning=FALSE}
source("scripts/chart_three.R")
chart_three <- make_chart_three(df)
chart_three
```

# Happiness Scores for the Lowest CO2 Per Capita Countries Bar Chart

```{r echo=FALSE, message=FALSE}
source("scripts/chart_one.R")
chart_co2_happiness_bar <- co2_happiness_bar(df)
chart_co2_happiness_bar
```

### What the chart attempts to seeks to express:

  What are the happiness scores of the 5 countries with 
  the lowest CO2 per capita in 2018? 


### What information this chart reveals: 
  
  The happiness scores for the 5 countries with the lowest CO2 per Capita 
  are 2.9 (Burundi), 3 (Central African Republic), 4.3 (Chad), 3.6 (Malawi), 
  and 5 (Somalia). This is interesting because the happiness scores of these 
  countries are in the bottom and middle of the range of happiness scores, thus
  the top five lowest CO2 per Capita countries are not countries with the highest
  happiness scores.
  
# CO2 Per Capita and Happiness Score Scatter Plot

```{r echo=FALSE, message=FALSE, warning = FALSE}
source("scripts/chart_two.R")
chart_co2_happiness_scatter <- co2_happiness_scatter(df)
chart_co2_happiness_scatter
```

### What the chart attempts to seeks to express: 
  
What relationship does the happiness score and CO2 per capita 
by country in 2018 have? Are countries with higher CO2 per capita happier or 
less happy?
  
### What information the chart reveals: 

The scatter plot suggests a positive relationship between CO2 per
capita and happiness score by country. Thus, in general, countries with higher
happiness scores also have higher co2 per capita emissions.
